Accelerating Neural Network Training using Arbitrary Precision Approximating Matrix Multiplication Algorithms

### 2021-03-03 latex/MPP21/paper.tex
added analysis of matmul accuracy data

Given these two contributions to the numerical error, $\lambda$ can be optimized to balance the effects based on parameters $\sigma$ and $\varphi$ (and $d$).
As shown by Bini, Lotti, and Romani \cite{BLR80}, the optimal $\lambda$ should be set to $\Theta(2^{-d/(\sigma+\varphi)})$.
Using this value of $\lambda$, the numerical error incurred by the algorithm will be bounded by $O(2^{-d\sigma/(\sigma+\varphi)})$.
Taking Bini's algorithm as an example, we have $\sigma=\varphi=1$, so the error is $O(2^{-d/2})$, or the square root of working precision.
Note that if multiple recursive steps are used, then $\varphi$ increases proportional to the number of steps, so straightforward optimization of $\lambda$ results in error of $O(2^{-d\sigma/(\sigma+s\varphi)})$ for $s$ recursive steps.
The parameters and minimum error for the algorithms we consider are presented in \cref{sec:APAprops}.
We show empirical error results for uniform random inputs of varying dimension in \cref{fig:accuracy}, as compared to the classical algorithm.
\GB{add more details on experiment here...}

@@

@@

\section{Neural Network Accuracy}

@@

\section{Neural Network Performance}

>>>>

Given these two contributions to the numerical error, $\lambda$ can be optimized to balance the effects based on parameters $\sigma$ and $\varphi$ (and $d$).
As shown by Bini, Lotti, and Romani \cite{BLR80}, the optimal $\lambda$ should be set to $\Theta(2^{-d/(\sigma+\varphi)})$.
Using this value of $\lambda$, the numerical error incurred by the algorithm will be bounded by $O(2^{-d\sigma/(\sigma+\varphi)})$.
Taking Bini's algorithm as an example, we have $\sigma=\varphi=1$, so the error is $O(2^{-d/2})$, or the square root of working precision.
Note that if multiple recursive steps are used, then $\varphi$ increases proportional to the number of steps, so straightforward optimization of $\lambda$ results in error of $O(2^{-d\sigma/(\sigma+s\varphi)})$ for $s$ recursive steps.
The parameters and minimum error for the algorithms we consider are presented in \cref{sec:APAprops}.

We show empirical error results for uniform random inputs of varying dimension in \cref{fig:accuracy}, as compared to the classical algorithm.
We measure the relative Frobenius norm error, or $\|C-\hat C\|_F/\|C\|$, where $\hat C$ is computed by each algorithm and $C$ is computed using the classical algorithm in double precision.
In order to choose the optimal $\lambda$ value for each algorithm, we tested the 5 powers of 2 closest to the theoretical optimal value and chose the best.

@@

Overall, we see little fluctuation of the error over matrix dimension, and the theoretical error bound is an upper bound on all empirical errors.
Note that the legend is ordered according to the error parameters, and the empirical error generally follows this ordering.
The two most accurate APA algorithms are the least recently discovered: Bini's $\dims{3,2,2}$ and Schonhage's $\dims{3,3,3;21}$.
Algorithms that offer more potential speedup tend to be less accurate.
Two algorithms with smaller error than expected are $\dims{5,5,5}$ and ${7,2,2}$, which is explained by the coefficients of those algorithms including fraction pre-factors.
While $\phi$ is computed using the exponents of the largest intermediate term of $\lambda$, the leading term of $\dims{5,5,5}$ has a constant of $1/4$, lessening its magnitude. 
We study the effects of the matrix multiplication error on neural network training and test accuracy in \cref{sec:NNacc}.

@@

\section{Neural Network Accuracy}
\label{sec:NNacc}

@@

\section{Neural Network Performance}
\label{sec:NNperf}
